---
title: "Death-Rate Estimation"
author: "Shreyas Gupta"
date: '2022-07-17'
output:
  html_document: default
  pdf_document: default
---
```{r, echo=TRUE}

library(readxl)
Altersverteilung <- as.data.frame(read_excel("Altersverteilung.xlsx", 
    sheet = "Fallzahlen"))

Todesfaelle <- as.matrix(read_excel("COVID-19_Todesfaelle.xlsx",
                                    sheet = "COVID_Todesfälle_KW_AG10" , ))

for(k in 1:119){
  for(l in 1:12){
    if(Todesfaelle[k,l] == "<4"){
       Todesfaelle[k,l] = 2 #set the mean of the uncertain 
                            #data as RKI only documented inequalities
    }
    Todesfaelle[k,l] = as.numeric(Todesfaelle[k,l])
  }
}  

```


Merge and order both data frames in order to receive homogeneous data set
of death ratios for each age group


```{r, echo=TRUE}

row.names(Altersverteilung) <- Altersverteilung[,1]
Altersverteilung <- as.matrix(Altersverteilung[-1,][,-1])
Altersverteilung <- Altersverteilung[,-(dim(Altersverteilung)[2]:
                                          (dim(Altersverteilung)[2]-1))]

X <- matrix(0, nrow = 10, ncol = 119)
rownames(X) <- rev(c("0-9","10-19","20-29","30-39",
                     "40-49","50-59","60-69","70-79","80-89","90+"))
colnames(X) <- colnames(Altersverteilung)[1:119]
X[1,] <- Altersverteilung[1,] 

```

Defining Y as Death Count and X as Infections:
Combining the age group to 10 year sequences: Adhering to the death-dataset.

```{r, echo=TRUE}
k <- 1
for (i in seq(2,19,2)){
  
  k = k+1
    
  X[k,] <- Altersverteilung[i,]+ Altersverteilung[(i+1),]
  
}
X<- X[rev(1:10),]
```

X is now the new Age distribution for Infections over the period of the pandemic.


```{r, echo=TRUE}

Y <- Todesfaelle[,-(1:2)]
Y <- apply(Y, 2, as.numeric)
Y <- t(Y)


rownames(Y) <- rownames(X)
colnames(Y) <- colnames(X)
```

Now, both, *Todesfaelle* and *Fallzahl*, are now the same mode and the same dimension. 


Bayesian Estimation
-------------------


Calculating the Ratios per Age-Group over all time at first.
Afterwards regardig the different time-periods 
and measures taken for each age group. 

```{r, echo=TRUE}
X_sum <- apply(X, 1, sum)
Y_sum <- apply(Y, 1, sum)

d_ratios <- Y_sum/X_sum
```


Now modeling the deaths as a Binomial Model regarding Infections:

 *Setting a Prior --> Beta Prior as Conjugate for Binomial
 *Naming the Likelihood

```{r, echo=TRUE}

plot(dbeta(x = seq(0,1, le=100),shape1 = 2, shape2 = 40))

alpha_0 <- 2
beta_0 <- 40

bayes_estimate <- (Y_sum + alpha_0)/(X_sum - Y_sum + beta_0)

comparison_frame <- data.frame(bayesestimate = bayes_estimate,
                               ratio = d_ratios, cases = X_sum) 

```

```{r, echo=TRUE}
require(ggplot2)
```

```{r, echo=TRUE}

ggplot(data = comparison_frame, aes(x = log(ratio), y = log(bayes_estimate),
                                    colour = log(cases)))+
  geom_point()+
  geom_abline(lty = 2, col= "red", lwd= .2)+
  geom_text(
    label=rownames(comparison_frame),
    check_overlap = F, 
    size = 3,
    nudge_x = -0.05, nudge_y = 0.15, 
  )


```

Now performing same estimation with different data window.

Covid Wave 1 
2020 Week 10 - Week 19

```{r, echo=TRUE}

X_1Wave <- X[,1:10]
Y_1Wave <- Y[,1:10]

X_sum1W <- apply(X_1Wave, 1, sum)
Y_sum1W <- apply(Y_1Wave, 1, sum)

d_ratios_1W <- Y_sum1W/X_sum1W


bayes_estimate_1W <- (Y_sum1W + alpha_0)/(X_sum1W - Y_sum1W + beta_0)
comparison_frame_1W <- data.frame(bayesestimate1W = bayes_estimate_1W,
                                  ratio1W = d_ratios_1W, cases = X_sum1W) 

```


Plotting differences

```{r, echo=TRUE}

ggplot(data = comparison_frame_1W, aes(x = log(ratio1W),
                                       y = log(bayesestimate1W),colour = log(cases)))+
  geom_point()+
  geom_abline(lty = 2, col= "red", lwd= .2)+
  geom_text(
    label=rownames(comparison_frame_1W),
    check_overlap = F, 
    size = 3,
    nudge_x = -0.05, nudge_y = 0.15, 
  )+
  xlim(-11,0)+
  ylim(-9,-1)

  

summary(comparison_frame_1W)
```

Now downphase: from 2020 Week 20 - Week 45

```{r, echo=TRUE}
X_1Down <- X[,11:36]
Y_1Down <- Y[,11:36]

X_sum1D <- apply(X_1Down, 1, sum)
Y_sum1D <- apply(Y_1Down, 1, sum)

d_ratios_1D <- Y_sum1D/X_sum1D


bayes_estimate_1D <-(Y_sum1D + alpha_0)/(X_sum1D - Y_sum1D + beta_0)
comparison_frame_1D <- data.frame(bayesestimate1D = bayes_estimate_1D,
                                  ratio1D = d_ratios_1D, cases = X_sum1D) 

```

```{r, echo=TRUE}
ggplot(data = comparison_frame_1D, aes(x = log(ratio1D), 
                                       y = log(bayesestimate1D),
                                       colour = log(cases)))+
  geom_point()+
  geom_abline(lty = 2, col= "red", lwd= .2)+
  geom_text(
    label=rownames(comparison_frame_1D),
    check_overlap = F, 
    size = 3,
    nudge_x = -0.05, nudge_y = 0.15, 
  )+
  xlim(-11,0)+
  ylim(-9,-1)

  

```

2 Wave 2020 Week 46 - 2021 Week 9

```{r, echo=TRUE}
X_2Wave <- X[,37:53]
Y_2Wave <- Y[,37:53]

X_sum2W <- apply(X_2Wave, 1, sum)
Y_sum2W <- apply(Y_2Wave, 1, sum)

d_ratios_2W <- Y_sum2W/X_sum2W


bayes_estimate_2W <- (Y_sum2W + alpha_0)/(X_sum2W - Y_sum2W + beta_0)
comparison_frame_2W <- data.frame(bayesestimate2W = bayes_estimate_2W,
                                  ratio2W = d_ratios_2W, cases = X_sum2W) 

```

Plot

```{r, echo=TRUE}
ggplot(data = comparison_frame_2W, aes(x = log(ratio2W), 
                                       y = log(bayesestimate2W),
                                       colour = log(cases)))+
  geom_point()+
  geom_abline(lty = 2, col= "red", lwd= .2)+
  geom_text(
    label=rownames(comparison_frame_2W),
    check_overlap = F, 
    size = 3,
    nudge_x = -0.05, nudge_y = 0.15, 
  )+
  xlim(-11,0)+
  ylim(-9,-1)

  
```


(MCMCM scheme: different Priors)
------------------------------


Using MCMC Metropolis Hastings Algorithm to implement uniform(beta) uninformative prior

Using the data defined in the first wave and evaluating posterior for age group 90+ :

```{r, echo=TRUE, eval=FALSE}

X_1Wave <- X[,1:10]; Y_1Wave <- Y[,1:10]
X_sum1W <- apply(X_1Wave, 1, sum); Y_sum1W <- apply(Y_1Wave, 1, sum)


proposal_frame <- data.frame(Infections = X_sum1W, Deaths = Y_sum1W)
```
Age Group data:
```{r, echo=TRUE, eval=FALSE}
group1_dframe <- c(rep(1,proposal_frame[10,2]),rep(0,proposal_frame[10,1]))
group2_dframe <- c(rep(1,proposal_frame[9,2]),rep(0,proposal_frame[9,1]))
group3_dframe <- c(rep(1,proposal_frame[8,2]),rep(0,proposal_frame[8,1]))
group4_dframe <- c(rep(1,proposal_frame[7,2]),rep(0,proposal_frame[7,1]))
group5_dframe <- c(rep(1,proposal_frame[6,2]),rep(0,proposal_frame[6,1]))
group6_dframe <- c(rep(1,proposal_frame[5,2]),rep(0,proposal_frame[5,1]))
group7_dframe <- c(rep(1,proposal_frame[4,2]),rep(0,proposal_frame[4,1]))
group8_dframe <- c(rep(1,proposal_frame[3,2]),rep(0,proposal_frame[3,1]))
group9_dframe <- c(rep(1,proposal_frame[2,2]),rep(0,proposal_frame[2,1]))
group10_dframe <- c(rep(1,proposal_frame[1,2]),rep(0,proposal_frame[1,1]))

```

Defining a function for all age groups
```{r, echo=TRUE, eval=FALSE}

wrapper_MH_posterior <- function(n.sims = 1000, alpha_0 = 2, beta_0 = 20, data_frame){
  
  theta_0 <- rbeta(n = 1, shape1 = alpha_0,shape2 = beta_0)
  theta_vec <- rep(0, n.sims +1) #create posterior vector
  theta_vec[1] <- theta_0 #initialize posterior vector
  
  for (t in 1:n.sims){
  
    theta_star <- rbeta(n = 1, shape1 = alpha_0,shape2 = beta_0)#proposal from prior    
    theta_t <- theta_vec[t] #previous accepted value for posterior  
    
    #Create MH-Ratio using log-likelihoods
    
    MH_Ratio <- sum(dbinom(data_frame, size = 1, log = T, prob = theta_star))/
      sum(dbinom(data_frame, size = 1, log = T, prob = theta_t))
    
    #sample with MH ratio
    
    mh_prob <- c(min(1,MH_Ratio), 1-min(1,MH_Ratio))
    
    theta_vec[t+1] <- sample(x = c(theta_star, theta_t),size = 1,prob = mh_prob)              
  }  
 
  hist(theta_vec, probability = T)
  #lines(dbeta(x = seq(0,1, le=1000), shape1 = alpha_0, shape2 = beta_0))
  
  plot(theta_vec, type = "l", main = "data_frame")
  #return(list(posterior = theta_vec,
  #quantiles = quantile(theta_vec, probs = c(0.1,0.9)), 
  #mean_posterior = mean(theta_vec)))
  
  return(mean(theta_vec))
}

#example

wrapper_MH_posterior(data_frame = group2_dframe, alpha_0 = 2 ,
                     beta_0 = 20, n.sims = 10000)

```


