---
title: "Function"
author: "Shreyas Gupta"
date: '2022-08-02'
output:
  pdf_document: default
  html_document: default
---

```{r}

Data <- read.csv(file = "Data.csv", header = TRUE, sep = ",", dec = ".")


n.states <- c("Schleswig-Holstein", "Hamburg","Niedersachsen","Bremen",
              "NRW","Hessen","Rheinland-Pfalz","Baden-Württemberg","Bayern","Saarland",
              "Berlin","Brandenburg","M-Vorpommern","Sachsen","Sachsen-Anhalt","Thüringen")

n.age <- c("A00-A04", "A05-A14", "A15-A34", "A35-A59", "A60-A79", "A80+")
gender <- c("M", "W", "unbekannt")

Data$Refdatum = as.Date(Data$Ref, format = "%Y-%m-%d")
```

The function takes input as:

data = which data, Bundeslaender = chceks if states are required,
Bundesland_Klasse = Which state is required, Landkreise = checks is counties are desired,
Landkreis_Klasse = selects county,Geschlechts_Klasse = Checks which gender, Alter_Klasse = Age Group,
alpha_0 = hyperparameter alpha, beta_0= hyperparameter beta,MH_algorithm = checks if running MH algo, 
Bayes_Estimate = checks if conjugation posterior should be computed,, n.sims = number of simulations,
time.stamp = whether a specific time frame is desired, Zeitpunkt = End data of desired frame, 
Zeitpunkt_Anfang = Start date of desired frame, beta = calculate for beta prior, uniform = Calculate
uniform prior if TRUE



```{r}
estimation <- function(data = Data, Bundeslaender = TRUE, Bundesland_Klasse, 
                       Landkreise = FALSE,Landkreis_Klasse,
                       Geschlechts_Klasse, Alter_Klasse, alpha_0, beta_0,
                       MH_algorithm = TRUE, Bayes_Estimate = TRUE, 
                       n.sims = 1000, time.stamp, Zeitpunkt, Zeitpunkt_Anfang,
                       beta = T, uniform = F, margins){
  
  #####set prior distribution
  
  expected_prior <- alpha_0/(alpha_0 + beta_0)
  
  ##prior is set as defined distribution which will be drawn from random
  
  if(Bundeslaender){
    frame <- data[(data$Bundesland == Bundesland_Klasse),]
  }
  if(Landkreise){
    frame <- data[(data$IdLandkreis == Landkreis_Klasse),]
  }
  frame <- frame[(frame$Geschlecht == Geschlechts_Klasse),]
  
  frame <- frame[(frame$Altersgruppe == Alter_Klasse),]
  
  frame <- frame[,c(5,10,11)]
  
  time_frame <- as.Date(frame[,1])  
  if(time.stamp){
    zeitpunkt_ende <- rev(time_frame)[1]
    zeitpunkt_anfang_daten <- time_frame[1]
    zeitpunkt_neu = as.Date(Zeitpunkt)
    zeitpunkt_anfang <- as.Date(Zeitpunkt_Anfang)
    
    if(sum(time_frame == zeitpunkt_neu, na.rm = T) == 0){
      while(sum(time_frame == zeitpunkt_neu, na.rm = T) == 0){
        zeitpunkt_neu <- zeitpunkt_neu +1
        zeitpunkt_neu <- as.Date(zeitpunkt_neu)
      }
    }  
    if(zeitpunkt_anfang < zeitpunkt_anfang_daten){
      while(zeitpunkt_anfang < zeitpunkt_anfang_daten){
        zeitpunkt_anfang <- zeitpunkt_anfang +1
        zeitpunkt_anfang <- as.Date(zeitpunkt_anfang)
      }
    }
    if(sum(time_frame == zeitpunkt_anfang, na.rm = T) == 0){
      while(sum(time_frame == zeitpunkt_anfang, na.rm = T) == 0){
        zeitpunkt_anfang <- zeitpunkt_anfang +1
        zeitpunkt_anfang <- as.Date(zeitpunkt_anfang)
      }
    }  
    start_date <- rev(which(time_frame %in% zeitpunkt_anfang))[1]
    end_date <- rev(which(time_frame %in% zeitpunkt_neu))[1]
    end_date_data <- rev(which(time_frame %in% zeitpunkt_ende))[1]
    
    
    if(zeitpunkt_ende < zeitpunkt_neu){
      frame <- frame[start_date:end_date_data ,]
    }
    else{
      frame <- frame[start_date:end_date ,]
    }
  }

  
  specific_data <- apply(frame[,2:3], 2, sum)
  ratio <- specific_data[2]/specific_data[1]
  
  ##build the data respective binomial distribution
  
  data_distribution <- c(rep(0, specific_data[1]), rep(1, specific_data[2]))
  
  ##### Bayes estimation -- no MH
  
  if(Bayes_Estimate){
    alpha_bayes <- specific_data[2] + alpha_0
    beta_bayes <- specific_data[1] + beta_0
    bayes_estimate <- alpha_bayes/(alpha_bayes + beta_bayes)
  }
  
  ####MCMC 
  theta_vec <- NULL
  
  if(MH_algorithm){
      
    if(beta){
      theta_0 <- rbeta(n = 1, shape1 = alpha_0, shape2 = beta_0)
    }
    if(uniform){
      theta_0 <- runif(0,0.5,n = 1)
    }  
    theta_vec <- rep(0, n.sims+1) #create posterior vector
    theta_vec[1] <- theta_0 #initialize posterior vector
    
    for (t in 1:n.sims){
      
      if(beta){
        theta_star <- rbeta(n = 1, shape1 = alpha_0, shape2 = beta_0)
      }
      if(uniform){
        theta_star <- runif(0,0.5,n = 1)
      }  
      
      #proposal from prior    
      theta_t <- theta_vec[t] #previous accepted value for posterior  
      
      #Create MH-Ratio using log-likelihoods and taking the inverse of each likelihood
      #to penalize higher negative values as outcomes with a smaller likelihood    
      
      MH_Ratio <- sum(dbinom(data_distribution, size = 1, log = T, prob = theta_t))/
        sum(dbinom(data_distribution, size = 1, log = T, prob = theta_star)) 
      
      #sample with MH ratio
      
      mh_prob <- c(min(1,MH_Ratio), 1-min(1,MH_Ratio))
      
      theta_vec[t+1] <- sample(x = c(theta_star, theta_t),size = 1,prob = mh_prob)              
    }  
    mh_estimate <- mean(theta_vec)
    quantiles <- quantile(theta_vec, probs = c(0.1,0.9))
    #par(mfrow = c(2,2))
    
    h <- hist(theta_vec[(0.3*n.sims):n.sims], plot = F, breaks = 10)
    cols <- c(0, 4, 0)
    k<- cols[findInterval(h$mids, quantile(theta_vec, probs = c(0.1,0.9)),
                          rightmost.closed=T, all.inside=F)+1]
    plot(h, col = k, main = paste("Histogram of posterior vector", Bundesland_Klasse), 
         xlab = "Values of Theta Vector", xlim = c(0,margins))
  }
  
  return(list(quantiles = quantiles, estimate = mean(theta_vec), ratio = ratio))
}

```


```{r, eval=TRUE}
estimation(Bundeslaender = T, Bundesland_Klasse = "Hamburg",
           Alter_Klasse = "A60-A79", alpha_0 = 4, beta_0 = 40,
           Zeitpunkt = "2020-05-06", Geschlechts_Klasse = "M", 
           time.stamp = T, n.sims = 1000, Zeitpunkt_Anfang = "2021-01-21",
           beta = F , uniform = T, margins = 0.5)
```

