---
title: "Political Evaluation CHhapter 6"
output:
  pdf_document: default
  html_document:
    df_print: paged
---


```{r}
Data <- read.csv(file = "Data.csv", header = TRUE, sep = ",", dec = ".")


n.states <- c("Schleswig-Holstein", "Hamburg","Niedersachsen","Bremen",
              "NRW","Hessen","Rheinland-Pfalz","Baden-Württemberg","Bayern","Saarland",
              "Berlin","Brandenburg","M-Vorpommern","Sachsen","Sachsen-Anhalt","Thüringen")

n.age <- c("A00-A04", "A05-A14", "A15-A34", "A35-A59", "A60-A79", "A80+")
gender <- c("M", "W", "unbekannt")

Data$Refdatum = as.Date(Data$Ref, format = "%Y-%m-%d")

```


```{r}
estimation <- function(data = Data, Bundeslaender = TRUE, Bundesland_Klasse, 
                       Landkreise = FALSE,Landkreis_Klasse,
                       Geschlechts_Klasse, Alter_Klasse, alpha_0, beta_0,
                       MH_algorithm = TRUE, Bayes_Estimate = TRUE, 
                       n.sims = 1000, time.stamp, Zeitpunkt, Zeitpunkt_Anfang){
  
  #####set prior distribution
  
  expected_prior <- alpha_0/(alpha_0 + beta_0)
  
  ##prior is set as defined distribution which will be drawn from random
  
  if(Bundeslaender){
    frame <- data[(data$Bundesland == Bundesland_Klasse),]
  }
  if(Landkreise){
    frame <- data[(data$IdLandkreis == Landkreis_Klasse),]
  }
  frame <- frame[(frame$Geschlecht == Geschlechts_Klasse),]
  
  frame <- frame[(frame$Altersgruppe == Alter_Klasse),]

  frame <- frame[,c(5,10,11)]
  
  time_frame <- as.Date(frame[,1])  
  if(time.stamp){
    zeitpunkt_ende <- rev(time_frame)[1]
    zeitpunkt_anfang_daten <- time_frame[1]
    zeitpunkt_neu = as.Date(Zeitpunkt)
    zeitpunkt_anfang <- as.Date(Zeitpunkt_Anfang)
    
    if(sum(time_frame == zeitpunkt_neu, na.rm = T) == 0){
      while(sum(time_frame == zeitpunkt_neu, na.rm = T) == 0){
          zeitpunkt_neu <- zeitpunkt_neu +1
          zeitpunkt_neu <- as.Date(zeitpunkt_neu)
      }
    }  
    if(zeitpunkt_anfang < zeitpunkt_anfang_daten){
      while(zeitpunkt_anfang < zeitpunkt_anfang_daten){
          zeitpunkt_anfang <- zeitpunkt_anfang +1
          zeitpunkt_anfang <- as.Date(zeitpunkt_anfang)
      }
    }
    if(sum(time_frame == zeitpunkt_anfang, na.rm = T) == 0){
      while(sum(time_frame == zeitpunkt_anfang, na.rm = T) == 0){
          zeitpunkt_anfang <- zeitpunkt_anfang +1
          zeitpunkt_anfang <- as.Date(zeitpunkt_anfang)
      }
    }  
    start_date <- rev(which(time_frame %in% zeitpunkt_anfang))[1]
    end_date <- rev(which(time_frame %in% zeitpunkt_neu))[1]
    end_date_data <- rev(which(time_frame %in% zeitpunkt_ende))[1]
    
  
    if(zeitpunkt_ende < zeitpunkt_neu){
      frame <- frame[start_date:end_date_data ,]
    }
    else{
    frame <- frame[start_date:end_date ,]
    }
  }
  else{
    frame <- frame[1:end_date_data ,]
  }
  
  specific_data <- apply(frame[,2:3], 2, sum)
  ratio <- specific_data[2]/specific_data[1]
  
  ##build the data respective binomial distribution
  
  data_distribution <- c(rep(0, specific_data[1]), rep(1, specific_data[2]))

  ##### Bayes estimation -- no MH
  
  if(Bayes_Estimate){
    alpha_bayes <- specific_data[2] + alpha_0
    beta_bayes <- specific_data[1] + beta_0
    bayes_estimate <- alpha_bayes/(alpha_bayes + beta_bayes)
  }

  ####MCMC 
  theta_vec <- NULL
    
  if(MH_algorithm){
  
    theta_0 <- rbeta(n = 1, shape1 = alpha_0, shape2 = beta_0)
    theta_vec <- rep(0, n.sims+1) #create posterior vector
    theta_vec[1] <- theta_0 #initialize posterior vector
    
    for (t in 1:n.sims){
      
      theta_star <- rbeta(n = 1, shape1 = alpha_0, shape2 = beta_0) #proposal from prior    
      theta_t <- theta_vec[t] #previous accepted value for posterior  
      
      #Create MH-Ratio using log-likelihoods and taking the inverse of each likelihood
      #to penalize higher negative values as outcomes with a smaller likelihood    
        
      MH_Ratio <- sum(dbinom(data_distribution, size = 1, log = T, prob = theta_t))/
        sum(dbinom(data_distribution, size = 1, log = T, prob = theta_star)) 
       
      #sample with MH ratio
      
      mh_prob <- c(min(1,MH_Ratio), 1-min(1,MH_Ratio))
      
      theta_vec[t+1] <- sample(x = c(theta_star, theta_t),size = 1,prob = mh_prob)              
    }  
    mh_estimate <- mean(theta_vec)
    quantiles <- quantile(theta_vec, probs = c(0.1,0.9))
    #par(mfrow = c(2,2))
 
    h <- hist(theta_vec[(0.3*n.sims):n.sims], plot = F, breaks = 10)
    cols <- c(0, 4, 0)
    k<- cols[findInterval(h$mids, quantile(theta_vec, probs = c(0.1,0.9)),
                          rightmost.closed=T, all.inside=F)+1]
    plot(h, col = k, main = paste("Histogram of posterior vector", Bundesland_Klasse), 
         xlab = "Values of Theta Vector", xlim = c(0,0.3))
  }
 
  return(list(quantiles = quantiles, estimate = mean(theta_vec), ratio = ratio))
}

```

## Function description

The function will find the nearest exact dates, if the dataset does not allow
the selection of that time point.Uses MH Algorithm. Returns quantiles and
estimate. 

Age Group = A80+
State = Rheinland-Pfalz

## Estimation during the second wave from 01.03.2020 - 06.05.2021

Pandemic was declared 11.03.2020

```{r}


W1 <- estimation(Bundesland_Klasse = "Rheinland-Pfalz", Geschlechts_Klasse = "M", 
           Alter_Klasse = "A80+", alpha_0 = 4, beta_0 = 40, time.stamp = TRUE,
           Zeitpunkt_Anfang = "2020-03-01",Zeitpunkt = "2020-05-06", n.sims = 1000)

```


## Estimation during the first down-phase from 07.05.2020 - 26.09.2020

```{r}

D1 <- estimation(Bundesland_Klasse = "Rheinland-Pfalz", Geschlechts_Klasse = "M", 
           Alter_Klasse = "A80+", alpha_0 = 4, beta_0 = 40, time.stamp = TRUE,
           Zeitpunkt_Anfang = "2020-05-07",Zeitpunkt = "2020-09-26", n.sims = 1000)


```

## Estimation during the second wave from 27.09.2020 - 27.02.2021

```{r}

W2 <- estimation(Bundesland_Klasse = "Rheinland-Pfalz", Geschlechts_Klasse = "M", 
           Alter_Klasse = "A80+", alpha_0 = 4, beta_0 = 40, time.stamp = TRUE,
           Zeitpunkt_Anfang = "2020-09-27",Zeitpunkt = "2021-02-27", n.sims = 1000)

```

## Estimation during the third wave from 28.02.2021 - 26.05.2021

```{r}

W3 <- estimation(Bundesland_Klasse = "Rheinland-Pfalz", Geschlechts_Klasse = "M", 
           Alter_Klasse = "A80+", alpha_0 = 4, beta_0 = 40, time.stamp = TRUE,
           Zeitpunkt_Anfang = "2021-02-28",Zeitpunkt = "2021-05-26", n.sims = 1000)



```


```{r}

estimates_MH <- c(W1$estimate, D1$estimate, W2$estimate, W3$estimate)
ratios <- c(W1$ratio, D1$ratio, W2$ratio, W3$ratio)
differene <- ratios - estimates_MH

comp_frame <- data.frame(estimates_MH, ratios, differene)

#xtable(comp_frame, digits = 4)

plot(1:4, estimates_MH, type = "l", col = "red", ylim = c(0.085, 0.12), axes = F, 
     xlab = "Phases", ylab = "Prob", main = "Course of Probs and Ratios")
legend(x = "bottomright", legend = c("Estimate MH", "Ratio"), lty = c(1,1), col = c("red", "blue"))
axis(1, at = 1:4, labels = c("Wave1","Down1","Wave2", "Wave3"))
axis(2)
lines(1:4, ratios, col = "blue")
segments(x0 = 1, y0 = estimates_MH[1], x1 = 1, y1 = ratios[1], lty = 2, lwd = .8, col = "gray")
segments(x0 = 2, y0 = estimates_MH[2], x1 = 2, y1 = ratios[2], lty = 2, lwd = .8, col = "gray")
segments(x0 = 3, y0 = estimates_MH[3], x1 = 3, y1 = ratios[3], lty = 2, lwd = .8, col = "gray")
segments(x0 = 4, y0 = estimates_MH[4], x1 = 4, y1 = ratios[4], lty = 2, lwd = .8, col = "gray")
```


